Learning Incoherent Subspaces: Classification via Incoherent Dictionary Learning.

### 2014-01-15 manuscript.tex
improved mcr plot

\maketitle
%
\begin{abstract}
In this article we present the supervised iterative projections and rotations (\Acronym{s-ipr}) algorithm, a method for learning discriminative incoherent subspaces from data. We derive \Acronym{s-ipr} as a supervised extension of our previously developed iterative projections and rotations (\Acronym{ipr}) algorithm for incoherent dictionary learning, and we employ it to learn incoherent sub-spaces that model signals belonging to different classes. We test our method as a feature transform for supervised classification, first by visualising transformed features from a synthetic dataset and from the `iris' dataset, then by using the resulting features in a classification experiment. While the visualisation results are promising, we find that \Acronym{s-ipr} generally performs worse than traditional and state-of-the-art techniques for supervised dimensionality reduction in terms of the misclassification ratio.
\end{abstract}
%
%\begin{keywords}
%Feature transforms, sparse approximation, dictionary learning, supervised classification.
%\end{keywords}
%
\section{Introduction: Classification And Feature Transforms}\label{sec:intro}
Supervised classification is one of the classic problems in machine learning where a system is designed to discriminate the category of an observed signal, having previously observed representative examples from the considered classes \cite{Duda1973Pa}.

@@

\subsection{Other related work}
The union of incoherent sub-spaces model proposed by Schnass and Vandergheynst \cite{Schnass2010A-} employes a very similar intuition to the one that inspired our proposed method, and models features belonging to different classes using incoherent subspaces.

Other methods for supervised dimensionality reduction include metric learning algorithms \cite{xing2002distance}, sufficient dimensionality reduction \cite{li1991sliced} and Bair's supervised principal components \cite{Bair06predictionby}.

@@

We employ the datasets detailed in Table \ref{tab:datasets}, and for each of them we evaluate the misclassification ratio, that is defined as the fraction of misclassified samples as a proportion of the total number of samples in the test set, averaged over the $5$ independent classification problems created by the stratified cross-validation protocol.
\begin{figure}
\centering
\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/fisheriris.pdf}}
\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/balance.pdf}}
\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/parkinsons.pdf}}
\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/sonar.pdf}}
%\subfigure{\includegraphics[width=.45\textwidth]{Code/Util/usps.pdf}}
\caption{\label{fig:class} Classification results.}
\end{figure}

@@

\section{Conclusion}\label{sec:end}
\subsection{Summary}
We have presented the \Acronym{s-ipr} algorithm for learning incoherent subspaces from data belonging to different categories. The encouraging experimental results obtained on the visualisation of the synthetic dataset and of a subset of features taken from the 'iris` dataset motivated us to test \Acronym{s-ipr} as a general method for feature transform to be used in classification problems. Unfortunately, we found that the performance of our proposed method on a group of datasets commonly used to benchmark classification methods is not competitive compared to traditional and state-of-the-art methods for feature transform.

@@

\subsection{Future work}
The main assumption made when using incoherent dictionary learning for classification is that high dimensional features are arranged onto lower-dimensional sub-spaces, and that features belonging to different classes can be modelled using different subspaces that are mutually incoherent. This assumption might be met by some datasets, but might not generally be satisfied by others. Understanding the general distribution of the features in a  dataset might be a necessary first step to inform a subsequent choice of algorithm, so that \Acronym{s-ipr} is only used in cases where its premise about the feature distribution is valid. This same argument holds for the whole class of linear models that comprises the dictionary learning model. Indeed, many feature transform techniques have equivalent kernelized versions to model non-linear data.

>>>>

\maketitle
%
\begin{abstract}
In this article we present the supervised iterative projections and rotations (\Acronym{s-ipr}) algorithm, a method for learning discriminative incoherent subspaces from data. We derive \Acronym{s-ipr} as a supervised extension of our previously proposed iterative projections and rotations (\Acronym{ipr}) algorithm for incoherent dictionary learning, and we employ it to learn incoherent sub-spaces that model signals belonging to different classes. We test our method as a feature transform for supervised classification, first by visualising transformed features from a synthetic dataset and from the `iris' dataset, then by using the resulting features in a classification experiment. While the visualisation results are promising, we find that \Acronym{s-ipr} generally performs worse than traditional and state-of-the-art techniques for supervised dimensionality reduction in terms of the misclassification ratio.
\end{abstract}
%
%\begin{keywords}
%Feature transforms, sparse approximation, dictionary learning, supervised classification.
%\end{keywords}
%
\section{Introduction: Classification And Feature Transforms}\label{sec:intro}
Supervised classification is one of the classic problems in machine learning where a system is designed to discriminate the category of an observed signal, having previously observed representative examples from the considered classes \cite{Duda1973Pa}.

@@

\subsection{Other related work}
The union of incoherent sub-spaces model proposed by Schnass and Vandergheynst \cite{Schnass2010A-} employes a very similar intuition to the one that inspired our proposed method, and models features belonging to different classes using incoherent subspaces. Other methods for supervised dimensionality reduction include metric learning algorithms \cite{xing2002distance}, sufficient dimensionality reduction \cite{li1991sliced} and Bair's supervised principal components \cite{Bair06predictionby}.

@@

We employ the datasets detailed in Table \ref{tab:datasets}, and for each of them we evaluate the misclassification ratio, that is defined as the fraction of misclassified samples as a proportion of the total number of samples in the test set, averaged over the $5$ independent classification problems created by the stratified cross-validation protocol.
\begin{figure}
\centering
\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/fisheriris.pdf}}
\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/balance.pdf}}
\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/parkinsons.pdf}}
\subfigure{\includegraphics[width=.48\textwidth]{Code/Util/sonar.pdf}}
%\subfigure{\includegraphics[width=.7\textwidth]{Code/Util/usps.pdf}}
\caption{\label{fig:class} Misclassification ratio as a function of the rank of the subspace employed during feature transforms for the datasets  `iris',`balance',`Parkinsons' and `sonar'}
\end{figure}

@@

\section{Conclusion}\label{sec:end}
\subsection{Summary}
We have presented the \Acronym{s-ipr} algorithm for learning incoherent subspaces from data belonging to different categories. The encouraging experimental results obtained on the visualisation of the synthetic dataset and of a subset of features taken from the 'iris` dataset motivated us to test \Acronym{s-ipr} as a general method for feature transform to be used in classification problems. Unfortunately, we found that the performance of our proposed method on a group of datasets commonly used to benchmark classification algorithms is only competitive compared to traditional and state-of-the-art methods for feature transform at high sub-space ranks.

@@

\subsection{Future work}
The main assumption made when using incoherent dictionary learning for classification is that high dimensional features are arranged onto lower-dimensional sub-spaces, and that features belonging to different classes can be modelled using different subspaces that are mutually incoherent. This assumption might be met by some datasets, but might not generally be satisfied by others. Understanding the general distribution of the features in a  dataset might be a necessary first step to inform a subsequent choice of algorithm, so that \Acronym{s-ipr} can be used in cases where its premise about the feature distribution is valid. This same argument holds for the whole class of linear models that comprises the dictionary learning model. Indeed, many feature transform techniques have equivalent \emph{kernelized} versions to model non-linear data.
